---
title: "Analysis of Health and Economic Impacts of Weather Events in the US"
output: 
      html_document:
            keep_md: true
---
## Synopsis

This quick analysis aims at defining which types of weather events have the strongest impacts on:
1. population health (measured in terms of fatalities and injuries) 
2. the economy (measured through property and crop damage)

## Data Processing

We start by loading the data into R. We then create specific data frames to look at particular values (sum of fatalities, injuries, property damage and crop damage)

```{r, echo=TRUE, cache = TRUE}
storm <- read.csv(bzfile("repdata_data_StormData.csv.bz2"))
```


## Analysis

```{r, echo=TRUE}
library(dplyr)
by_ev <- group_by(storm, EVTYPE)
test <- summarise(by_ev, sum(FATALITIES))
names(test)[2] <- "SUM_FATAL"
fatal <- test[test$SUM_FATAL != 0, ]
fatal <- fatal[order(fatal$SUM_FATAL, decreasing = TRUE),]

top <- fatal$SUM_FATAL[1:10]
names(top) <- fatal$EVTYPE[1:10]
barplot(top, cex.names=0.7, las=2, main="Total fatalities per event type - 1950-2011")
```



## Results

This rough analysis shows that *tornados* are the most harmful weather events in the US in terms of casualties.
However, the first step to getting a more accurate picture would be to clean up the categorization of event types in the dataset (many distinct EVTYPE values for a single actual event type, such as *thunderstorm wind*)
